---
author: Logan Johnson
title: "'plyr' to d'plyr'"
categories: Split-apply-combine
date: 2023-02-16
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Prompt:

The `plyr` package has by now been replaced by other, even faster
packages, but the idea of *Split, apply, combine* is as relevant as
ever.

Read the paper [The Split-Apply-Combine Strategy for Data
Analysis](https://www.jstatsoft.org/article/view/v040i01) by Hadley
Wickham.

Write a blog post addressing the following questions:

1.  The R code for the split-apply-combine paper is posted with the
    paper. Pick one of the examples demonstrating `plyr` functionality
    (such as `dlply` or `ddply`, ...) and rewrite the example using
    functionality from the package `dplyr`. Make sure that your example
    works and the results are identical.
    
## Example from plyr package
    
```{r}
library(plyr)

baseball <- baseball

head(baseball[c("id","year")], n = 10)

baseball_1 <- ddply(baseball, .(id), transform, 
  cyear = year - min(year) + 1)

head(baseball_1[c("id","year","cyear")], n = 10)

tail(baseball_1[c("id","year","cyear")], n = 10)

```


## Remade with dplyr package

```{r}
library(dplyr)

baseball <- baseball

head(baseball[c("id","year")], n = 10)

baseball_2 <- baseball %>% 
  arrange(id) %>% 
  group_by(id) %>% 
  mutate(cyear = year-min(year) + 1)

head(baseball_2[c("id","year","cyear")], n = 10)

tail(baseball_2[c("id","year","cyear")], n = 10)

```

2.  Which (base R) functions do you know that support the
    split-apply-combine strategy? In your opinion, are these sufficient?
    State why or why not?
    
- I'm not super familiar with base R functions in the split-apply-combine area. I know of apply and lapply, and I have heard of some of the other functions. I honestly have not used them before. I do feel like they take more work to achieve the similar function of dplyr. The coding and naming of the base R functions is confusing to me. I do think that there are probably use cases for these functions compared with the dplyr package. Overall, I do not think that the base R functions for the split-apply-combine strategy are sufficient because they are not as user accessible for normal people. It would take more time and understanding of R to be able to achieve a similar output in base R compared to dplyr.

3. The ChatGPT response to the prompt "Describe the split-apply-combine paradigm" is given below. Generally, the chatbot's answers are decent. ChatGPT does have problems getting details right. Which details would you change in the response? 

    
>    The "split-apply-combine" paradigm is a data analysis strategy used in statistics and data science to analyze grouped data. It consists of three steps:
>
>    Split: The data is first split into groups based on one or more variables of interest.
>
>    Apply: A specific operation or function is applied to each group of data, usually for the purpose of aggregating, summarizing, or transforming the data within each group.
>
> Combine: The results of the operation applied to each group are then combined and returned as a single output.
>
> The split-apply-combine paradigm is particularly useful in cases where the data cannot be analyzed as a whole and instead must be analyzed in smaller, more manageable chunks. This allows for more efficient data processing and improved scalability. The paradigm is commonly implemented using tools such as the "groupby" function in the R and Python programming languages, and the "dplyr" library in R.

- I think I would change the last part of the response. I don't think that the data is necessarily split into chunks but rather managed in a more efficient manner that allows for a function or modification to be applied throughout the desired parts of the data.


You can write your answers directly the `README.Rmd` file. 
Make sure that the file knits (you will need to activate any packages your use in your code chunks with a call to `library(xxx)`, where xxx is the name of the package, such as `plyr` ).
Commit your changes and push to your repo;  add any files in the `README_files` directory to your repository. 


