---
author: "AR"
title: "plyr to dplyr"
category: Split-apply-combine
date: 2022-02-16
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Prompt:

The `plyr` package has by now been replaced by other, even faster
packages, but the idea of *Split, apply, combine* is as relevant as
ever.

Read the paper [The Split-Apply-Combine Strategy for Data
Analysis](https://www.jstatsoft.org/article/view/v040i01) by Hadley
Wickham.

Write a blog post addressing the following questions:

1.  The R code for the split-apply-combine paper is posted with the
    paper. Pick one of the examples demonstrating `plyr` functionality
    (such as `dlply` or `ddply`, ...) and rewrite the example using
    functionality from the package `dplyr`. Make sure that your example
    works and the results are identical.
    

```{r}
library(plyr)
library(dplyr)

baseball <- ddply(baseball, .(id), transform,
  cyear = year - min(year) + 1)


#using dplyr
baseball%>%
  group_by(id)%>%
  mutate(cyear = year - min(year) + 1)


```
I chose to replace the ddply function. ddply splits data frames by variables, which can be done in dplyr with group_by. mutate creates new columns that are functions of existing variables.

2.  Which (base R) functions do you know that support the
    split-apply-combine strategy? In your opinion, are these sufficient?
    State why or why not?

-	split()
-	lapply()
-	rbind()

are all functions in base R. I think they work well, but maybe not as efficiently as plyr. However, I do believe they are sufficient. I use rbind and lapply() all the time. 


3. The ChatGPT response to the prompt "Describe the split-apply-combine paradigm" is given below. Generally, the chatbot's answers are decent. ChatGPT does have problems getting details right. Which details would you change in the response? 

First thing is, I wouldn’t say “analyze grouped data”. To me, it isn’t incorrect, just not great wording. It also missed that split-apply-combine can be used in base R and plyr.  Also, it missed that it is a paradigm and therefor has many applications outside of just data cleaning and manipulation. As Hadley Wickham's paper says, you can use it in modeling and visualization as well. 
    


