---
title: Errors and warnings in packages
author: Tyler Wiederich
date: "2023-03-30"
categories: "Errors and warnings in packages"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- badges: start -->
[![Frontmatter check](../../actions/workflows/check-yaml.yaml/badge.svg)](../../actions/workflows/check-yaml.yaml)
<!-- badges: end -->

## Prompt:

Fix one of the problems in one of our community packages at `https://github.com/Stat585-at-ISU/lab-3-all-all-for-one-and-one-for-all`, and write about it. 



The error that I fixed is not actually one that R detected. 
When I was finding errors in the various packages, I noticed that the file names when downloading the pdf file from Team6 attached "NULL" to the end of the file name. 
The output came from the `get_pd_presslog` function.
This was easily fixed by changing the `cat` function into `paste`. 

```{r}
#Original line
print(cat('Some text:', 'user input'))

#New line
print(paste('Some text:', 'user input'))
```

The problem with these types of errors is that they can sometimes go unnoticed when writing code. 

There was one other error that I fixed. 
The function arguments for the url and save locations were assigned immediately inside the function, so any user input for these arguments would be ignored.

```{r, eval=FALSE}
#Original format
function(arg1, arg2){
  arg1 <- 'Some default url'
  arg2 <- 'Some default save location'
  #Rest of function here
}

#New format
function(arg1 = 'Some default url', arg2 = 'Some default save location'){
  #Rest of function here
}
```

Again, this is an error that R would not alert users of since it computationally does not produce any warnings or errors. 
In this case, the danger would be old files with the same name would be replaced when running the function again. 
The url for the source would not change, but it is better to put the default into the function argument rather than writing over the argument in the code.








