---
author: "Harun Celik"
title: "Split-Apply-Combine Post"
categories: "All that legal stuff..."
date: "2023-03-09"
output:
  html_document: default
  github_document: default
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->

[![Frontmatter
check](../../actions/workflows/check-yaml.yaml/badge.svg)](../../actions/workflows/check-yaml.yaml)
[![Render
rmarkdown](../../actions/workflows/render-rmarkdown.yaml/badge.svg)](../../actions/workflows/render-rmarkdown.yaml)

<!-- badges: end -->

## Prompt:

The `DESCRIPTION` file of a package contains the package's meta
information. Most of the fields in this file are quite straight forward:
author, version number, and a short package description. When you call
`library(help="<package name>")` for package `<package name>` you can
see the contents of the `DESCRIPTION` file for that package (and some
parts of the `NAMESPACE` file).

Read through Colin Fay's (short) book on [Licensing
R](https://thinkr-open.github.io/licensing-r/)

Write a blog post addressing the following questions:

### Under what license does R operate? What is the license for ggplot2?

Running `license()` shows us that R operates under the GNU General
Public License for both Version 2 and Version 3. Running `?ggplot2`
shows that the copyright holder is with RStudio and the CRAN repository
for the package indicates that gpplot2 operates under the MIT + file
LICENSE.

### What are the dependencies of the package you made?

The following are the dependencies of our package based on our manual
addition of the `usethis` package:

-   dplyr - MIT + file LICENSE (formerly GPL license)

-   lubridate - GNU GPL (\>=2) (it would be interesting to see what
    happens to this now that it will roll with tidyverse)

-   readr - MIT + file LICENSE

-   purrr - MIT + file LICENSE

-   stringr - MIT + file LICENSE

-   tabulizer - MIT + file LICENSE

We can check all of the dependencies with a custom function that filters
through CRAN database since most of the these are on CRAN.

```{r}
library(dplyr)
pkgdb <- tools::CRAN_package_db()
```

The author of the book suggested using `{pkgnet}` but I had trouble
understanding exactly what the custom function was doing there so I just
filtered through the `pkgdb` data and output the items as a list for
legibility. This will show the dependencies, imports and suggests for
each package listed above.

```{r}
get_deps <- function(pkg) {
  pkgdb %>%
    select(Package, Depends, Imports, Suggests) %>%
    filter(Package == pkg) %>%
    as.list()
}

get_deps("dplyr")
get_deps("lubridate")
get_deps("purrr")
get_deps("readr")
get_deps("stringr")

# Not surprisingly tabulizer doesn't return any values since it is not on CRAN
get_deps("tabulizer")
```

For `tabulizer`, we will need to get this information from the github
repository for the package in the DESCRIPTION file. The only way I've thought of 
being able to access this is through the URL but I don't think this is modular enough
for all packages installed on github with differing remote urls. The `{desc}` package
seems only helpful for manipulating already documented DESCRIPTION files, while the `{remotes}`
package seems to provide options for installation, I could not figure out how to
retrieve the DESCRIPTION data for a particular package :( 

Read through [chapter 5.2 of Licensing
R](https://thinkr-open.github.io/licensing-r/practical.html#dependencies-exploration).
Combine and adjust the code pieces to create the dependencies for a
package that is on github rather than CRAN. (Don't forget about the
package `desc`.) What dependencies does the package have that you
created for Lab-3? Which licenses are involved? Are all of these
dependencies required? Elaborate.

Under which license should this blog post be published? Make sure that
you are using an appropriate license. (You can change the current
licenses with the help of commands from the `use_license_XXX` family in
the `usethis` package)

This blog currently is operating under the Creative Commons Legal Code CC0 1.0 
Universal license. It does use RMarkdown which operates under the GNU General
Public License and the GNU recommends that the license be distributed in the form
of a GPL. To change this, we can use the `{usethis}` package.

```{r}
library("usethis")
usethis::use_gpl_license(version = 3)
```
This package is now licensed by GPL Version 4. This can be checked in the `License.md`
file.

### Can I publish an R package on CRAN under an MIT license when I have a dependency on a package that is licensed under GPL-3?

When confronted with the above question, ChatGPT gave two -- quite
different -- responses, as shown below. The main point in both responses
is 'your package may be considered derivative work'. At the specific
problem of the package developed for Lab 3, what is your opinion about
potential copyright infringements when releasing the package with an MIT
license? Make sure to take the relevant passage in Colin Fay's book
under consideration.

    The legal implications of publishing an R package on CRAN under an MIT license while having a dependency on a package that is licensed under GPL-3 are complex and can vary depending on the specific circumstances. It is always advisable to seek legal advice from a licensed attorney if you have any doubts or concerns about licensing issues.

    However, in general, it is possible to publish an R package on CRAN under an MIT license when you have a dependency on a package that is licensed under GPL-3. This is because the MIT license is a permissive open-source license that allows for more flexibility in how the software can be used and distributed, and it is generally considered compatible with the GPL-3 license.

    That being said, if your package depends on a GPL-3-licensed package, your package may be considered a derivative work and therefore subject to the terms of the GPL-3 license. This means that if you distribute your package, you may need to also distribute the source code for your package under the GPL-3 license or a compatible license, and you may need to ensure that the requirements of the GPL-3 license are met for the entire distribution, including any dependencies.

    It is also worth noting that the R community has its own guidelines and recommendations for package licensing, and it is important to comply with these guidelines if you want to publish your package on CRAN. The CRAN Repository Policy states that packages should have a "permissive" or "GPL-compatible" license, but does not explicitly require that packages use the GPL license or any other specific license.

    In summary, while it is possible to publish an R package on CRAN under an MIT license while having a dependency on a package that is licensed under GPL-3, it is important to understand the implications of the different licenses and to ensure that you comply with any applicable guidelines and recommendations from the R community.

After harping a bit on copyright infringement in general, ChatGPT
changed its response to

    Technically, you can publish an R package on CRAN under an MIT license when you have a dependency on a package that is licensed under GPL-3, but doing so would not be compliant with the terms of the GPL-3 license.

    The GPL-3 license is a strong copyleft license that requires derivative works to be released under the same license. If your package depends on a GPL-3-licensed package, your package may be considered a derivative work and therefore must be licensed under the GPL-3 license or a compatible license.

    The MIT license, on the other hand, is a permissive license that allows for more flexibility in how the software is used and distributed. However, it is not compatible with the GPL-3 license, and therefore cannot be used for derivative works of GPL-3-licensed software.

    Publishing an R package on CRAN under an MIT license when you have a dependency on a package that is licensed under GPL-3 would be a licensing violation, and could potentially result in legal action or other consequences. It is important to comply with the licensing requirements of all dependencies in your package to avoid any legal issues.

The 'harping' questions:

-   Is it legal to publish an R package on CRAN under an MIT license
    when I have a dependency on a package that is licensed under GPL-3?

-   Is it illegal to publish an R package on CRAN under an MIT license
    when I have a dependency on a package that is licensed under GPL-3?

-   Is it a copyright infringement to publish an R package on CRAN under
    an MIT license when I have a dependency on a package that is
    licensed under GPL-3?

-   Can I publish an R package on CRAN under an MIT license when I have
    a dependency on a package that is licensed under GPL-3 without
    infringing copyright?

## Submission

Write answers to the questions directly into the file `README.Rmd`. Push
the blog post to your blog-7 repo. Make sure that all of the checks are
passing.
