---
author: Caio dos Santos
date: 2023-01-26
category: "Asking Good Questions"
title: Asking Good Questions
---

## THE POST

### [Calculate distance between latitude and longitude points?](https://stackoverflow.com/questions/75228511/calculate-distance-between-latitude-and-longitude-points/75228753#75228753)

It seemed like the Stack Overflow user [Jamie](https://stackoverflow.com/users/20152268/jamie) was having some trouble calculating the distance between a hospital in Fresno, CA (119.85 °W, 36.84° N) and the home location of several patients.

Jamie had a data frame with the geographical coordinates of city in which the hospital patients lived and wanted to calculate the geographical distance that each patient traveled.

Jamie provided the geographical coordinates of the hospital

```{r eval=FALSE}
Latitude: 36.840

Longitude: -119.850
```

In addition, Jamie also provided the an example data frame

```{r eval=FALSE}
structure(list(major_city = c("Mountain View", "Watsonville", "Honolulu", "Los Altos", "Morgan Hill", "Fulton", "Oak Grove", "Port Kent", "Bedford", "San Jose"), 

latitude = c(37.39, 36.9, 21.31, 37.36, 37.1, 43.3, 36.69, 44.526, NA, 37.27),

longitude = c(-122.07, -121.7, -157.85, -122.15, -121.7, -76.4, -87.44, -73.409, NA, -121.84)), 
row.names = c(NA, -10L), class = c("data.table", "data.frame"), 
.internal.selfref = <pointer: 0x7f98af80bae0>, index = integer(0))

```

I provided an answer using the *apply* function. The idea was to apply a function rowwise which returned the distance from the hospital to the home location of each patient.

```{r, eval=FALSE}
df <- data.frame(major_city = c("Mountain View","Watsonville", "Honolulu", "Los Altos", "Morgan Hill", "Fulton", "Oak Grove", "Port Kent", "Bedford", "San Jose"),
                 latitude = c(37.39, 36.9, 21.31, 37.36, 37.1, 43.3, 36.69, 44.526, NA, 37.27),
                 longitude = c(-122.07, -121.7, -157.85, -122.15, -121.7, -76.4, -87.44, -73.409, NA, -121.84))

df$distance_m <- apply(df, MARGIN = 1, function(x) {
  geosphere::distGeo(as.numeric(c(x['longitude'], x['latitude'])),
                     c(-119.85, 36.84)) 
  })

df$distance_km <- df$distance_m / 1000

```

The ouput provided the distance for all rows in the data frame

```{r, eval=FALSE}
      major_city latitude longitude distance_m distance_km
1  Mountain View   37.390  -122.070   206527.4    206.5274
2    Watsonville   36.900  -121.700   165083.2    165.0832
3       Honolulu   21.310  -157.850  4045683.3   4045.6833
4      Los Altos   37.360  -122.150   212440.1    212.4401
5    Morgan Hill   37.100  -121.700   167241.2    167.2412
6         Fulton   43.300   -76.400  3731976.3   3731.9763
7      Oak Grove   36.690   -87.440  2879678.0   2879.6780
8      Port Kent   44.526   -73.409  3962114.4   3962.1144
9        Bedford       NA        NA         NA          NA
10      San Jose   37.270  -121.840   183321.3    183.3213
```

## MY EXPERIENCE

This was my first time answering questions on Stack Overflow. This question in specific was good because the poster provided a minimal reproducible example. When I was looking for questions to answer, I saw several questions that could not be answered because they did not provide examples. Overall, it was a pleasant experience, being able to provide answers to a community that has greatly helped me when I taking my first steps towards learning R.
