---
title: 'Lab Assignment #2: Getting that data'
author: "Heike Hofmann"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, echo = TRUE)
```
# Outline

- `pdf` files are for text, right?

- Looking under the hood ...



The final deliverable is again a self-contained RMarkdown file and a dataset. 

---

# Lab organization

1. As last time, there is a github classroom link. Select the invite with your team's number. 

2. Hofmann, don't forget the Zoom breakout rooms. Join the room that matches your team number.

3. Final deliverable: submit a link to your repository in Canvas (This will show me that you are done working on your project).

4. Create a folder `data` in the repository. Assume that all of the data files used are inside this folder, similar to the code used in this Rmarkdown file. 



---

# First item: talk!

Just as with lab #1, you are asked to create one file (README.Rmd) with your results. 

All team members are supposed to contribute ... last time you might have realized that it is a pretty big mess if everybody writes into the README file all at the same time.

This time: figure out a plan before you start working. Include your plan as a table in the README.Rmd file for lab #2. 

---

# Our Goal for this lab

- read a pdf file into an R session

- go through cleaning steps to turn the pdf into usable information

- save the resulting code and data files


---

# Service calls to the Ames police 

are collected and published each day for a time interval of the past four days as 'Press log'

https://www.cityofames.org/government/departments-divisions-i-z/police/news-and-updates

available under 

https://data.city.ames.ia.us/publicinformation/PressLog.pdf

---

# Download the Press log file

and inspect it ... 

```{r}
presslog_url <- "https://data.city.ames.ia.us/publicinformation/PressLog.pdf"
download.file(presslog_url, destfile = "PressLog.pdf", mode = "wb")
```

---

# Reading a pdf file

The package `tabulizer` provides bindings (i.e. wrapper functionalities) for the java library Tabula PDF Table Extractor.

Our first goal for the lab is to install the `tabulizer` package.

Go to https://docs.ropensci.org/tabulizer/ and follow the  directions for installing the package

---

# `extract_tables`

The function `extract_tables` returns a list object. 

Describe the structure of this list:

- What does the length of the list correspond to? 

- What objects are the list's items? 

- What are the dimensions of each of the list's items? 

```{r eval=FALSE}
library(tabulizer)
plog <- extract_tables(file="PressLog.pdf")

str(plog)
```
---

# Transform the `plog` object into a clean data set

Data set of service calls: 

- format of object is data frame or tibble

- columns are variables with appropriate names

- each row corresponds to one incident

- all variables have the correct format

- save the result as a csv file with the name "presslog-yyyymmdd.csv', where yyyymmdd is the date of first reported incident the file 

---

# Helper files and code

Data set of call codes: 

- data frame (or tibble) with two columns named 'Abbreviation' and 'Description'

- save the result as a csv file with the name "call-codes.csv"

Save the code (just the R script, not an Rmd) in a file called "process_presslog_yyyymmdd.R"


---

# How robust is your code?

The folder 'Presslogs' contains service call files from a couple of other days. 

Process these files as well.
Do you need to make any adjustments to your code? Keep track of those adjustments.


Save data and code for these presslogs as: 

  - "presslog-yyyymmdd.csv" 
  - "process_presslog_yyyymmdd.R"

---

# Deliverables (1) or (2)

## Deliverable (1)

Create a markdown table with a description of the structure of your repository (i.e. which files are where, and what is included).
The structure is not going to be graded itself, just your description of the structure. 
You will find that good structures are easier to describe :)

Combine all of your service call data sets into one, draw a barchart of the number of service calls by day, colour by service log file.


## Deliverable (2)

In case things went horribly wrong, and you could not get things to work, describe what you did, which error you encountered, and what strategies you tried before you gave up/ the time for the lab was up. 

---

# Submit

<br>

Push your changes to the repository, make sure that the **README.Rmd file knits** on your machines, push the **README.md** file as well. 

Make sure to add all datasets and all code files to the repository.

Upload a link to your repository to Canvas. Only one submission per team is required. 
Finishing touches can be made until Monday, Feb 20, 11:59 pm.

<br>
<br>
<br>
And you are done!

