---
author: "Valerie Han"
title: "The Split-Apply-Combine Strategy"
categories: Split-apply-combine
date: 2023-02-16
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Prompt:

The `plyr` package has by now been replaced by other, even faster
packages, but the idea of *Split, apply, combine* is as relevant as
ever.

Read the paper [The Split-Apply-Combine Strategy for Data
Analysis](https://www.jstatsoft.org/article/view/v040i01) by Hadley
Wickham.

Write a blog post addressing the following questions:

1.  The R code for the split-apply-combine paper is posted with the
    paper. Pick one of the examples demonstrating `plyr` functionality
    (such as `dlply` or `ddply`, ...) and rewrite the example using
    functionality from the package `dplyr`. Make sure that your example
    works and the results are identical.
    
    


```{r paper example}
library(plyr)
baseball1 <- ddply(baseball, .(id), transform, 
  cyear = year - min(year) + 1)

```


```{r}
library(dplyr)
baseball2 <- baseball %>%
  group_by(id) %>%
  mutate(cyear = year - min(year) + 1) %>%
  arrange(id) %>%
  as.data.frame()

identical(baseball1, baseball2)
```




2.  Which (base R) functions do you know that support the
    split-apply-combine strategy? In your opinion, are these sufficient?
    State why or why not?


With a dataframe `df` with a column `col1`, you can subset by value val1 of `col1` with `df[df$col1 == val1, ]`. You could then do whatever procedure you wanted on that subsetted dataframe. There's also the `apply` functions in base R (or a for loop), which would allow you to apply the same procedure to multiple values and combine the values. However, I don't believe the base R functions are sufficient. The split-apply-combine strategy is so common in data analysis that it shouldn't take so many (not super readable) lines of code to accomplish it. The functions in `dplyr` are much more intuitive to code with and understand, reducing the cognitive load.



3. The ChatGPT response to the prompt "Describe the split-apply-combine paradigm" is given below. Generally, the chatbot's answers are decent. ChatGPT does have problems getting details right. Which details would you change in the response? 

    
>    The "split-apply-combine" paradigm is a data analysis strategy used in statistics and data science to analyze grouped data. It consists of three steps:
>
>    Split: The data is first split into groups based on one or more variables of interest.
>
>    Apply: A specific operation or function is applied to each group of data, usually for the purpose of aggregating, summarizing, or transforming the data within each group.
>
> Combine: The results of the operation applied to each group are then combined and returned as a single output.
>
> The split-apply-combine paradigm is particularly useful in cases where the data cannot be analyzed as a whole and instead must be analyzed in smaller, more manageable chunks. This allows for more efficient data processing and improved scalability. The paradigm is commonly implemented using tools such as the "groupby" function in the R and Python programming languages, and the "dplyr" library in R.


One important thing that the answer doesn't mention is that the operations on each group must be assumed to be independent. In addition, the last paragraph has a few misleading details. The data usually can be analyzed as a whole, but analyzing by group gives us different information than analyzing as a whole. In addition, the paradigm is not so much useful in runtime efficiency/scalability but rather in reducing the cognitive load for certain common data analysis tasks (which can increase efficiency in writing/understanding code). The last sentence would also be more accurate as: The paradigm is  implemented using tools such as the `group_by` and `groupby` functions in the `dplyr` and `pandas` libraries in the R and Python programming languages, respectively.



You can write your answers directly the `README.Rmd` file. 
Make sure that the file knits (you will need to activate any packages your use in your code chunks with a call to `library(xxx)`, where xxx is the name of the package, such as `plyr` ).
Commit your changes and push to your repo;  add any files in the `README_files` directory to your repository. 






