---
author: Parvin Mohammadiarvejeh
date: 2023-01-26
categories: "Asking Good Questions"
title: Asking Good Questions
---

## Prompt:

Asking good questions is a valuable skill to have - asking questions in an online setting is both easier and harder than asking questions in person: we can prepare to ask a question but we are also expected to prepare.
The links posted here give some advice on how to ask good questions:

- stackoverflow's [Asking a good question](http://stackoverflow.com/help/how-to-ask)

- R's [Posting guidelines](https://www.r-project.org/posting-guide.html)

- [minimal complete verifiable example](https://stackoverflow.com/help/mcve), [minimal reproducible example](https://www.tidyverse.org/help/)

Follow these links and read through the advice given, then

1. **Pick at least one question from stackoverflow or the R help and answer it.**

Write a blog post answering the following questions: 

2. **Document which question you answered (link to your answer).**

3. **Relate your experience of answering the question to your reading. **
 

<!--Go to [https://github.com/Stat585-at-ISU/blog](https://github.com/Stat585-at-ISU/blog) for instructions about how to prepare and submit your blog post.-->
Instructions to follow.


# Parvin's work

1 & 2. I picked this question, link: (https://stackoverflow.com/questions/75225812/how-to-identify-rows-where-observations-are-identical-in-two-different-columns-i)

My stackoverflow account name: StarWiiin



In the following R chunk, I prepared my solution with code:
```{r}
# Create a data including two columns: start_id, end_id

start_id = c("one", "two", "three", "four", "five", "six", "seven")
end_id = c("one", "three", "two", "four", "seven", "six", "five")


df = data.frame(start_id, end_id)
df

```



```{r}
# create a new column which is True if start_id equals end_id, otherwise it is False.
library(dplyr)
df = df %>% 
  mutate(status= if_else(.$start_id == .$end_id, 'True', 'False'))
df
```



3. I tried to answer breifly but comprehensive. Since the question was related to R, I created a dataframe based on the writer explanations and I gave the solution exactly by solving the example. Also, I got the output myseld and I checked it to be sure it is true. So, my solution is easily reproducible and verified. 


















